Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-22 paper/paper.tex
addressed Ewa's comments

We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2). In
this model, virtual machine (VM) instances are requested on-demand. The VMs are
billed by the hour, with partial hours being rounded up. Although there are
multiple VM types with different amounts of CPU, memory, disk space, and I/O,
for this paper we focus on a single VM type because we assume that for most
applications there will typically be one VM type with the best price/performance
ratio for the application~\cite{Juve2009}. We also assume that a submitted task
has exclusive access to a VM instance and that there is no preemption.

@@

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
(top) and {\em large budget} (bottom).}
\label{fig:number-complete-pareto}
\end{figure*}

@@

\begin{figure*}[t]  
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{ Total work completed expressed as the sum of runtimes (in hours) of
tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows}
\label{fig:total-time}
\end{figure*}

@@

\begin{figure*}[htb] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calculated by dividing the sum of
task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows.}
\label{fig:cost}
\end{figure*}

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values for each application ensemble. The budget values
selected correspond to two extreme cases: {\em small budget} (top) and {\em
large budget} (bottom). The small budgets permit the algorithms to complete 
only a small number of workflows from the ensemble, while the larger budgets 
enable the algorithms to finish almost all the workflows. In three of five 
cases the SPSS algorithm performs better for tighter constraints. When the 
budget is low, or the deadline is short, the static procedure is able to more 
efficiently pack workflow tasks onto limited resources. This is the case for 
the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
budgets with short deadlines. The exceptions are the Montage and CyberShake 
ensembles,  which consist of many fine-grained tasks. In these applications, 
as well for the other ensembles with large budgets and deadlines, the simple 
dynamic scheduling techniques perform equally well or better. The reason for 
this behavior is that the large number of tasks guarantees that resources 
are almost never idle regardless of how the tasks are scheduled. We also 
observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
This means that for the workflows we tested the simple admission procedure based 
on estimation of workflow cost does not degrade the solution, but rather in 
many cases it enables larger workflows that would lead to budget overrun to
be rejected. Thus it can save the space for smaller workflows to complete.

@@

\begin{figure}[htb] \centering
\includegraphics[width=0.7\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}\\
(a) ensemble of Pareto-distributed workflows
\includegraphics[width=0.7\columnwidth]{figures/constant-budget-SIPHT-n-1000-0-dagh5-50h4m0.pdf}\\
(b) ensemble of workflows of equal size \caption{Comparison of number of
workflows completed for a given deadline, plotted versus different budget
amounts.}
\label{fig:budgets}
\end{figure}

@@

Similar results were achieved for ensembles consisting of workflows of equal
size. A summary of these results is shown in Table~\ref{tab:num-dags-equal}.
Again, we can see that the static SPSS algorithm performed better for coarse-grained
workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
are smaller since the workflows have equal size, therefore rejecting a particular
workflow does not make it any easier for the algorithms to complete any other 
workflow.

Fig.~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
perspective. It helps answer the question: Given a deadline, how many workflows
can be finished when we increase the budget? In Fig.~\ref{fig:budgets}(a) the
growth is super-linear, which results from the Pareto-like distribution of 
workflow sizes in the ensemble (see Fig.~\ref{fig:ensemble-distribution}). 
For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
the growth is linear.

>>>>

We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2). In
this model, virtual machine (VM) instances are requested on-demand. The VMs are
billed by the hour, with partial hours being rounded up. Although there are
multiple VM types with different amounts of CPU, memory, disk space, and I/O,
for this paper we focus on a single VM type because we assume that for most
applications there will typically be one VM type with the best price/performance
ratio for the application~\cite{Juve2009}. We also assume that a submitted task
has exclusive access to a VM instance and that there is no preemption. For this
study we assumed that there are no delays in VM start and terminate requests.

@@

\begin{figure*}[t] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
(top) and {\em large budget} (bottom). The ranges on $y$ axis are chosen to 
emphasize the differences the performance of the algorithms.}
\label{fig:number-complete-pareto}
\end{figure*}

@@

\begin{figure*}[t]  
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{ Total work completed expressed as the sum of runtimes (in hours) of
tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
algorithms on ensembles of 100 Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
emphasize the differences between the performance of the algorithms.}
\label{fig:total-time}
\end{figure*}

@@

\begin{figure*}[htb] 
\centering
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
\caption{Effective computing cost in \$ per hour calculated by dividing the sum of
task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
emphasize the differences the performance of the algorithms.}
\label{fig:cost}
\end{figure*}

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values for each application ensemble. The budget values
selected correspond to two extreme cases: {\em small budget} (top) and {\em
large budget} (bottom). The small budgets permit the algorithms to complete 
only a small number of workflows from the ensemble, while the larger budgets 
enable the algorithms to finish almost all the workflows. The rangesIn three of
five cases the SPSS algorithm performs better for tighter constraints. When the 
budget is low, or the deadline is short, the static procedure is able to more 
efficiently pack workflow tasks onto limited resources. This is the case for 
the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
budgets with short deadlines. The exceptions are the Montage and CyberShake 
ensembles,  which consist of many fine-grained tasks. In these applications, 
as well for the other ensembles with large budgets and deadlines, the simple 
dynamic scheduling techniques perform equally well or better. The reason for 
this behavior is that the large number of tasks guarantees that resources 
are almost never idle regardless of how the tasks are scheduled. We also 
observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
This means that for the workflows we tested the simple admission procedure based 
on estimation of workflow cost does not degrade the solution, but rather in 
many cases it enables larger workflows that would lead to budget overrun to
be rejected. Thus it can save the space for smaller workflows to complete.

@@

\begin{figure}[htb] \centering
\includegraphics[width=0.7\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h4m0.pdf}\\
(a) ensemble of Pareto-distributed workflows
\includegraphics[width=0.7\columnwidth]{figures/constant-budget-SIPHT-n-1000-0-dagh5-50h4m0.pdf}\\
(b) ensemble of workflows of equal size \caption{Comparison of number of
workflows completed for a given deadline, plotted versus different budget
amounts.}
\label{fig:budgets}
\end{figure}

@@

Similar results were achieved for ensembles consisting of workflows of equal
size set to 1000. A summary of these results is shown in
Table~\ref{tab:num-dags-equal}. Again, we can see that the static SPSS algorithm performed better for coarse-grained
workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
are smaller since the workflows have equal size, therefore rejecting a particular
workflow does not make it any easier for the algorithms to complete any other 
workflow.

Fig.~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
perspective. It helps answer the question: Given a deadline, how many workflows
can be finished when we increase the budget? In Fig.~\ref{fig:budgets}(a) the
growth is super-linear, which results from the Pareto-like distribution of 
workflow sizes in the ensemble (see Fig.~\ref{fig:ensemble-distribution}). 
For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
the growth is linear and the costs are higher, since all the worklfows in the
test consisted of 1000 tasks.
